---
title: "Analysis of Measured MFI in Untreated Tissue Sample"
output: github_document
---

## Analysis of MFI values of Tissue Samples

This is an R Markdown format used for publishing markdown documents to GitHub. When you click the **Knit** button all R code chunks are run and a markdown file (.md) suitable for publishing to GitHub is generated.

#install necessary packages
```{r setup, include=FALSE}

install.packages("ggplot2")
install.packages("tidyverse")
install.packages("brms")
install.packages("conflicted")
install.packages("dplyr")
install.packages("magrittr")
install.packages("conflicted")
install.packages("readr")

library(ggplot2)
library(tidyverse)
library(tidyr)
library(brms)
library(conflicted)
library(dplyr)
filter <- dplyr::filter
library(magrittr) #for pipeline %>% used in ggplot
library(readr)
install.packages("gridExtra")
library(gridExtra)

```

## Download 2 files for MFI Inner (Caseum) and Outer (Cuff) Location

```{files for Caseum and Cuff region}
B_0274_inner_ROI <- read_csv("B_0274_inner.csv", 
                             col_types = cols_only(`Maximum intensity_0` = col_guess(), 
                                                   `Maximum intensity_1` = col_guess(), 
                                                   `Mean Intensity_0` = col_guess(), 
                                                   `Mean Intensity_1` = col_guess(), 
                                                   `Size in pixels` = col_guess(), `Total Intensity_0` = col_guess(), 
                                                   `Total Intensity_1` = col_guess(), 
                                                   `Minimum intensity in neighborhood_0` = col_guess(), 
                                                   `Minimum intensity in neighborhood_1` = col_guess(),
                                                   object_id = col_guess()))

B_0274_outer_ROI <- read_csv("B_0274_outer.csv", 
                             col_types = cols_only(`Maximum intensity_0` = col_guess(), 
                                                   `Maximum intensity_1` = col_guess(), 
                                                   `Mean Intensity_0` = col_guess(), 
                                                   `Mean Intensity_1` = col_guess(), 
                                                   `Size in pixels` = col_guess(), `Total Intensity_0` = col_guess(), 
                                                   `Total Intensity_1` = col_guess(), 
                                                   `Minimum intensity in neighborhood_0` = col_guess(), 
                                                   `Minimum intensity in neighborhood_1` = col_guess(),
                                                   object_id = col_guess()))

```
All files will be downloaded to current directory as two individual datasets

## Bind separate dataset

Rename column names to bind datasets:

```{Rename Columns}
BI <- data.frame(Object_ID = B_0274_inner_ROI$object_id , Treatment = "Untreated" ,  
                              Location = "Inner" , 
                              Pixel_Size = B_0274_inner_ROI$`Size in pixels`,
                              Max_Red = B_0274_inner_ROI$`Maximum intensity_0`, 
                              Max_Green = B_0274_inner_ROI$`Maximum intensity_1`, 
                              MFI_Red = B_0274_inner_ROI$`Mean Intensity_0`, 
                              MFI_Green = B_0274_inner_ROI$`Mean Intensity_1`,
                              Nei_Red = B_0274_inner_ROI$`Minimum intensity in neighborhood_0`,
                              Nei_Green = B_0274_inner_ROI$`Minimum intensity in neighborhood_1`,
                              Total_Red = B_0274_inner_ROI$`Total Intensity_0`,
                              Total_Green = B_0274_inner_ROI$`Total Intensity_1`)



BO <- data.frame(Object_ID = B_0274_outer_ROI$object_id , Treatment = "Untreated" ,  
                               Location = "Outer" , 
                               Pixel_Size = B_0274_outer_ROI$`Size in pixels` ,
                               Max_Red = B_0274_outer_ROI$`Maximum intensity_0`, 
                               Max_Green = B_0274_outer_ROI$`Maximum intensity_1`, 
                               MFI_Red = B_0274_outer_ROI$`Mean Intensity_0`, 
                               MFI_Green = B_0274_outer_ROI$`Mean Intensity_1`,
                               Nei_Red = B_0274_outer_ROI$`Minimum intensity in neighborhood_0`,
                               Nei_Green = B_0274_outer_ROI$`Minimum intensity in neighborhood_1`,
                               Total_Red = B_0274_outer_ROI$`Total Intensity_0`,
                               Total_Green = B_0274_outer_ROI$`Total Intensity_1`)


```

## Combine datasets

Combine dataframes BI and BO and create new columns for background corrected MFI values:

```{Combine Factors}
Pivot_Data<- rbind (BI, BO)
Pivot_Data$Red <- Pivot_Data$MFI_Red - Pivot_Data$Nei_Red
Pivot_Data$Green <- Pivot_Data$MFI_Green - Pivot_Data$Nei_Green
Pivot_Data$Ratio <- Pivot_Data$Red / Pivot_Data$Green

#create dataset with original values for MFI
pivot_data_origina_background <- Pivot_Data %>%
  pivot_longer(
    cols = c('Red', 'Green'),
    names_to = "Channel",
    values_to = "MFI")
#to log transform the values for graphs to remove outliers
Pivot_Log_Data <- Pivot_Data
Pivot_Log_Data$Red <- log10(Pivot_Log_Data$Red)
Pivot_Log_Data$Green <- log10(Pivot_Log_Data$Green)

summary(Pivot_Data$Ratio)
Pivot_Data$Log_Ratio <- log10(Pivot_Data$Ratio)
summary(Pivot_Data$Log_Ratio)

summary(Pivot_Data$Ratio)

dat_back_3 <- Pivot_Data %>%
  pivot_longer(
    cols = c('Red', 'Green'),
    names_to = "Channel",
    values_to = "MFI")  
```
## filter ratio values based on location for new datasets for analysis

```{Separate location for analysis}
dat_back_3_Inner <- filter(dat_back_3, Location == "Inner")
dat_back_3_Outer <- filter(dat_back_3, Location == "Outer")

quantile(dat_back_3_Inner$Ratio, 0.25)
quantile(dat_back_3_Inner$Ratio, 0.50)
quantile(dat_back_3_Inner$Ratio, 0.75)
IQR(dat_back_3_Inner$Ratio)

IQR(dat_back_3_Outer$Ratio)
quantile(dat_back_3_Outer$Ratio, 0.25)
quantile(dat_back_3_Outer$Ratio, 0.50)
quantile(dat_back_3_Outer$Ratio, 0.75)

```


#Info on test selected for comparing differences between mean values

Kruskal wallis one-way ANOVA method for comparing k independent samples
For groups identically distributed (null hypothesis?) But also used to compare
median values. The test assumes that obser. in each group are identically and
independently distributed apart from location. This form of testing is selected where
the assumptions for parametric anova are not met.

```{sd}
#One-Way non-parametric ANOVA (Kruskal Wallis Test)
str(tree_1$Location) # make sure location is a factor
aggregate(Log_Ratio~Location, data = dat_back_3, sd)
```
Location                Log_Ratio
1    Inner 0.6175149843218231682229
2    Outer 0.8049127793920733964228
```{mean}
aggregate(Log_Ratio~Location, data = dat_back_3, mean)
```
Location                  Log_Ratio
1    Inner -0.64518047927759303306061
2    Outer  0.06216522540183311296147
```{variance}
aggregate(Log_Ratio~Location, data = dat_back_3, var)
```
Location                Log_Ratio
1    Inner 0.3813247558619815014147
2    Outer 0.6478845824286726395869

Tests for log transformed ratio values based on location
```{KW Test}
      kruskal.test(Log_Ratio~Location, data = dat_back_3)
```

#Kruskal-Wallis rank sum test
data:  Log_Ratio by Location
Kruskal-Wallis chi-squared = 51409.69115646242426, df = 1, p-value < 2.2204460492503131e-16

Results show that the p-value was significant for dominance of median values
between groups. The null hypothesis can be rejected
Pairwise comparisons using Wilcoxon rank sum test

```{pairwise wilcox test}
      pairwise.wilcox.test(dat_back_3$Log_Ratio, dat_back_3$Location, paired = FALSE, p.adj = "BH")
```

data:  tree_1$Log_Ratio and tree_1$Location 
Inner                 
Outer < 2.22044604925031e-16
P value adjustment method: BH

Pairwise wilcox test uses the pairwise command to test every sample to
everyother sample to calculate pairwise compairisons between group levels
with corrections for multiple testing.


#Kruskal Wallis test Inner location Pre vs 23s
boxplots tell us that the MFI values need to be log transformed to account for outliers
```{boxplot, echo = FALSE}
boxplot(Log_Ratio~Location, data = dat_back_3_Inner)
```

```{pairwise wilcox test}
kruskal.test(Log_Ratio_Location, data = dat_back_3_Inner)
```
Kruskal-Wallis rank sum test

data:  MFI by Channel
Kruskal-Wallis chi-squared = 83967.41080970132316, df = 1, p-value <
  2.2204460492503131e-16

Results show that there are differences in dominance of medians of MFI channels 23s and PrerRNA
```{pairwise wilcox test}
pairwise.wilcox.test(dat_back_3_Inner$MFI, Tree_Inner$Channel, paired = FALSE, p.adj = "BH")
```
Pairwise comparisons using Wilcoxon rank sum test 
data:  Tree_Inner$MFI and Tree_Inner$Channel 
Green                 
Red < 2.22044604925031e-16
P value adjustment method: BH 

Results show that there are significant differences between groups 23s and pre for MFI

#Testing for variance within groups
```{mean}
aggregate(MFI~Channel, data = dat_back_3_Inner, mean)
```
#Channel                     MFI
1   Green 3.505382695053870456547
2     Red 2.860202215776277423487
```{sd}
aggregate(MFI~Channel, data = dat_back_3_Inner, sd)
```
```{variance}
aggregate(MFI~Channel, data = dat_back_3_Inner, var)
```
#Channel                       MFI
1   Green 0.06849946342499749207811
2     Red 0.28568441451431980659947
boxplot(MFI~Channel, data = Tree_Inner)


#Kruskal Wallis test Outer location Pre vs 23s
Boxplots tell us that the MFI values need to be log transformed to account for outliers
```{kruskal-wallis}
kruskal.test(MFI~Channel, data = dat_back_3_Outer)
```
Kruskal-Wallis rank sum test
data:  MFI by Channel
Kruskal-Wallis chi-squared = 446.1596899552567379, df = 1, p-value <
  2.2204460492503131e-16


Results show that there are differences in dominance of medians of MFI channels 23s and PrerRNA
#pairwise wilcox test
```{pairwise wilcox test}
pairwise.wilcox.test(dat_back_3_Outer$MFI, dat_back_3_Outer$Channel, paired = FALSE, p.adj = "BH")
```
Pairwise comparisons using Wilcoxon rank sum test 
data:  dat_back_3_Outer-MFI dat_back_3_Outer$Channel 
Green                 
Red < 2.22044604925031e-16
P value adjustment method: BH 

Results show that there are significant differences between groups 23s and pre for MFI

#Extended Figure 3c Test for variance

#Testing for variance within groups
```{pairwise wilcox test}
aggregate(MFI~Channel, data = dat_back_3_Outer, mean)
```
#Channel                    MFI
1   Green 3.48042161673112104836
2     Red 3.54258684213295405030
```{pairwise wilcox test}
aggregate(MFI~Channel, data = dat_back_3_Outer, sd)
```
```{pairwise wilcox test}
aggregate(MFI~Channel, data = dat_back_3_Outer, var)
```
#Channel                      MFI
1   Green 0.1289192666483613691764
2     Red 0.4341223453724966652167

```{r}
#Kruskal Wallice Test for MFI values

#Testing for variance
aggregate(Red~Location, data = dat_back_3, var)

Location       Red
1   Caseum  15389649
2     Cuff 104985177

#Testing for variance in green based on location
aggregate(Green~Location, data = dat_back_3, var)

Location   Green
1   Caseum 4671665
2     Cuff 9515047

#Testing median values within locations for MFI Pre-rRNA

kruskal.test(Red~Location, data = dat_back_3)
Kruskal-Wallis rank sum test

data:  Red by Location
Kruskal-Wallis chi-squared = 30211, df = 1, p-value < 2.2e-16

##test for difference insw MFI values based on location with
non.04 <- lm(Red~Location, data=dat_back_3)
summary(non.04)
table(cat_3$Location)
# Testing median values within locations for MFI Green
kruskal.test(Green~Location, data = dat_back_3)

> kruskal.test(Green~Location, data = dat_back_3)

Kruskal-Wallis rank sum test

data:  Green by Location
Kruskal-Wallis chi-squared = 0.25117, df = 1, p-value = 0.6163

```

# F-test for variance 
F-test to compare two variances
A F-test was conducted in order to compare variance between two samples, in this case inner versus outer location. Results show that we can reject the null hypothesis that there is no variability between groups. A p-value of 2.2 e-16 well below the significance level of 0.05, for MFI values for both Pre-rRNA and 23S- rRNA, indicates that there is a significant difference in measured MFI values based on location.
# separate df into two group : cuff and caseum for location. 
# set values that will be used for models
# attach(df)
# names(df)
# F_Inner <- var.test(Log_Ratio ~ Location, data = df,
# alternative = “two.sided”)

F test to compare two variances
data: Log_Ratio by Location
F = 0.14095, num df = 126572, denom df = 38304, p-value < 2.2e-16 alternative hypothesis: true ratio of variances is not equal to 1
95 percent confidence interval:
0.1386878 0.1432446 sample estimates: ratio of variances
0.1409528


# F_outer <- var.test(Log_Ratio ~ Location, data = df,
# alternative = “two.sided")
F test to compare two variances
data: Log_Ratio by Location
F = 0.4865, num df = 126572, denom df = 38304, p-value < 2.2e-16 alternative hypothesis: true ratio of variances is not equal to 1
95 percent confidence interval:
0.4786864 0.4944145 sample estimates: ratio of variances
0.4865042

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
